---
title: "Data Visualization Session"
subtitle: "Usage of unhcthemes package"
author: "Cedric Vidonne"
institute: "IPDA/GDS"
date: "`r format(Sys.Date(),  '%d %B %Y')`"
output:
  unhcrdown::html_slides:
    self_contained: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Objectives

**About today:**

- Basic of the grammar of graphics and [**`ggplot2`**](https://ggplot2.tidyverse.org/index.html)
- Introduction to the [**`unhcrthemes`**](https://vidonne.github.io/unhcrthemes/) package
- Best practices to create **UNHCR branded** visuals in R, with examples
- What else, dynamic report with the [**`unhcrdown`**](https://github.com/vidonne/unhcrdown)

**Not today:**

- Data import: [`readr`](https://readr.tidyverse.org/), [`readxl`](https://readxl.tidyverse.org/), etc.
- Data manipulation: [`dplyr`](https://dplyr.tidyverse.org/), [`tidyr`](https://tidyr.tidyverse.org/), etc.
- R programming: [R for Data Science](https://r4ds.had.co.nz/), [Advanced R Programming](https://adv-r.hadley.nz/), etc.

--

See the `xaringan::moon_reader()` package documentation<sup>1</sup> and examples<sup>2</sup> for more information regarding usage.


.footnote[
[1] https://github.com/yihui/xaringan

[2] https://slides.yihui.org/xaringan/
]


---
class: center, middle

# Section title

### An interesting new Section

---

# Basic style

The templates comes with predefined styles for inline text, such as *italic*, **bold** or <u>underline text</u>.

It also includes the requires styles for common elements such as:

.pull-left[
**Headers:**

## Heading 2

### Heading 3

**Lists:**

* item 1
* item 2
    1. sub-item 2.1
    2. sub-item 2.2
]

.pull-right[

**Or quotes:**

> It is further proof that refugee movements and the broader issue of migration of populations ... is a global challenge that cannot be confined to a few countries.
>
> Filippo Grandi

]

---
class: inverse, center, middle

# Section title - Dark

### An interesting new Section

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r}
knitr::kable(head(iris), format = 'html')
```

---
background-image: url(`r xaringan:::karl`)
background-position: 50% 50%
class: bottom, inverse, hide-logo

### Or a full page picture with no logo

---

# Reveal

Here's a new slide that show piece by piece

--

* One

--

* Two

--

* Three

---
class: inverse, center, middle

# Thank you

### Goodbye slide
