---
title: "unhcRstyle Workshop Plot Exercise"
author: "Cedric Vidonne"
date: "March 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>",
  dev = 'ragg_png',
  dpi = 300,
  warning = FALSE
)
```

## Install and load required packages

```{r load-packages}
# Uncomment the lines below if you don't have the packages installed
# install.packages(c('tidyverse', 'readxl', 'scales'))
# remotes::install_github('unhcr/unhcrdatapackage')
# remotes::install_github('unhcr-web/unhcRstyle')
library(tidyverse)
library(unhcrdatapackage)
library(unhcRstyle)
library(readxl)
library(scales)
```

## Load and clean required datasets

```{r load-data}
# End of year populatiuon from UNHCR Data finder
popdata <- unhcrdatapackage::end_year_population_totals

# 4W data from Colombia
data_4W <- read_excel("data/4W_Publico_Consolidado_septiembre.xlsx")

# Manipulate to get same stage as R basic workshop
data_4W %>%
  mutate(Municipio = str_to_lower(Municipio),
         Municipio = str_replace_all(Municipio, "á", "a"),
         Municipio = str_replace_all(Municipio, "í", "i"),
         Municipio = str_replace_all(Municipio, "ó", "o"),
         Municipio = str_replace_all(Municipio, "é", "e"),
         Municipio = str_replace_all(Municipio, "ú", "u")) %>%
  distinct(Municipio)

data_4W %>%
  mutate(Departamento = str_to_lower(Departamento)) %>%
  distinct(Departamento)

length(union(data_4W$`Socio principal`, data_4W$`Socio implementador`))

data_4W %>%
  filter(Estado == "Completada") %>%
  summarise(sum(Cantidad, na.rm = TRUE))

data_4W %>%
  filter(Estado == "Completada") %>%
  summarise(sum(`Beneficiarios mensuales`, na.rm = TRUE))

```

## Bar chart

**Exercise 1:** Create a bar chart of the number of activity per sector using the unhcr_theme and playing with the visual aspect.

There are two ways of representing bar chart with geom_bar or geom_col. The main difference is that geom_bar comes with a stat of count as default. This can be overwrite with `stats = "identity"` (see results below for examples).

**Advanced** sort it by number of activity and color it by part of the RMRP or not.

```{r bar-ex}
# Place your code here

```

**Results:**

```{r bar-sector}
# Example using geom_bar and simply counting the number of row per sector.
bar_sector <- ggplot(data = data_4W) +
  aes(x = Sector) +
  geom_bar(fill = unhcr_blue) +
  coord_flip() +
  unhcr_theme(grid = "Xx") +
  labs(title = "Number of activities per sector",
       caption = "Source: 4W Colombia",
       x = NULL,
       y = "Number of activity")

bar_sector

bar_sector_sorted <- ggplot(data = data_4W) +
  aes(x = forcats::fct_rev(forcats::fct_infreq(Sector))) + # Creates order by number of appearance and reverse it to have the maximum at the top.
  geom_bar(fill = unhcr_blue) +
  coord_flip() +
  unhcr_theme(grid = "Xx") +
  labs(title = "Number of activities per sector",
       caption = "Source: 4W Colombia",
       x = NULL,
       y = "Number of activity")

bar_sector_sorted

bar_sector_sorted_fill <- ggplot(data = data_4W) +
  aes(x = forcats::fct_rev(forcats::fct_infreq(Sector))) + 
  geom_bar(aes(fill = `Parte RMRP`)) + # Create fill colors
  coord_flip() +
  unhcr_theme(grid = "Xx", legend_title = TRUE) + # Add title to the legend
  labs(title = "Number of activities per sector",
       caption = "Source: 4W Colombia",
       x = NULL,
       y = "Number of activity") +
  scale_fill_manual(values = usecol(pal = c(unhcr_red, unhcr_blue))) # Set colors to the fill

bar_sector_sorted_fill
```


```{r col-sector}
# Example using geom_col. Need to create the count in the data first
data_4W_counted <- data_4W %>% 
  count(Sector, `Parte RMRP`, name = "count") %>% 
  group_by(Sector) %>% 
  mutate(total_count = sum(count)) %>% 
  ungroup()

col_sector <- ggplot(data = data_4W_counted) +
  aes(x = Sector, y = count) +
  geom_col(fill = unhcr_blue) + # or geom_bar(stat = "identity")
  coord_flip() +
  unhcr_theme(grid = "Xx") +
  labs(title = "Number of activities per sector",
       caption = "Source: 4W Colombia",
       x = NULL,
       y = "Number of activity")

col_sector

col_sector_sorted <- ggplot(data = data_4W_counted) +
  aes(x = reorder(Sector, total_count), y = count) +
  geom_col(fill = unhcr_blue) +
  coord_flip() +
  unhcr_theme(grid = "Xx") +
  labs(title = "Number of activities per sector",
       caption = "Source: 4W Colombia",
       x = NULL,
       y = "Number of activity")

col_sector_sorted

col_sector_sorted_fill <- ggplot(data = data_4W_counted) +
  aes(x = reorder(Sector, total_count), y = count) +
  geom_col(aes(fill = `Parte RMRP`)) +
  coord_flip() +
  unhcr_theme(grid = "Xx", legend_title = TRUE) +
  labs(title = "Number of activities per sector",
       caption = "Source: 4W Colombia",
       x = NULL,
       y = "Number of activity")  +
  scale_fill_manual(values = usecol(pal = c(unhcr_red, unhcr_blue)))

col_sector_sorted_fill

```

## Line chart

**Exercise 2:** Create a line chart of the number of IDPs in Colombia from 2000 to 2019 using the unhcr_theme and playing with the visual aspect.

**Advanced** add more population type and label directly the lines.

```{r line-ex}
# Place your code here

```

**Results:**

```{r line-result}
# Work on the data
displaced_col <- popdata %>% 
  filter(CountryAsylumCode == "COL",
         Year >= 2000) %>% 
  group_by(Year) %>% 
  summarise(IDP = sum(IDP, na.rm = TRUE))

# Plot the data
ggplot(data = displaced_col) +
  aes(x = Year, y=IDP, group = 1) +
  geom_line(color = unhcr_green, size = 1) +
  unhcr_theme(grid = "Yy") +
  labs(title = "Total number of IDPs in Colombia",
       subtitle = "Evolution of the IDP population 2000-2019",
       caption = "Source: UNHCR Data Finder") +
  scale_y_continuous(labels = scales::label_number_si(),
                     name = "Number of IDP") +
  scale_x_discrete(breaks = c(2000,2005, 2010, 2015, 2019), name = "")
```

## Pie chart

**Exercise 3:** Create a pie chart of the proportion of Modalidad

**Advanced** add direct labels and improve order of slices.

```{r line-ex}
# Place your code here

```

**Results:**

```{r pie-result}
# Work on the data
modalidad <- data_4W %>% 
  group_by(Modalidad) %>% 
  summarise(count = n())

# Plot the data
ggplot(data = modalidad) +
  aes(x = "", y = count, fill = Modalidad) + 
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start=0) +
  theme_void() +
  scale_fill_manual(values = usecol(c(unhcr_pal_grey[2], unhcr_pal_blue[3]))) +
  labs(title = "Proportion of CASH assistance",
       caption = "Source: 4W Colombia")
```
